---
title: "Election"
format: html
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(primer.data)
library(brms)
library(tidybayes)
library(gtsummary)
```

# Overview
Using information about candidates for governor from 1945 to 2012, we seek to find the relationship between death age of Preceptor and whether or not he wins the mayoral election in Newton, Massachusetts. Modern medicine has increased the overall lifespan of all candidates regardless if they won or lost. We modeled the age of death as a sum of multiple variables multiplied by their coefficients: election result, how much they won by, age during the election, political party, and sex. We are 95% confident that Preceptor will live an extra 3 to 14 years. Our estimate is 8.6 years.

# Question
How old will Preceptor live to if he wins the election for mayer of Newton?

# Wisdom

## Preceptor Table
A perfect table. One row for each mayer, columns are state, lived_after. Covariates are state, sex, election_age, party. 

## EDA
```{r}
#| label: EDA

x <-
governors |>
  mutate(treatment = ifelse(win_margin > 0, "win", "lose")) |>
  filter(abs(win_margin) < 5)

x |>
  mutate(name = paste(first_name, last_name, sep = " ")) |>
  select(name, election_age, state, year, lived_after) |>
  ggplot(aes(x = election_age, y = lived_after)) +
  geom_point()
```


## Validity
Relationship between columns in preceptor table is the same for the data. One problem is governor is not a mayer

# Justice

## Population table
table with preceptor table, data, and greater population

## Stability
People are living longer overtime because of better healthcare

## Representativeness
Massachusetts represents more than Newton

## Unconfoundedness
it's a predictive model

not random because the people volunteered to run


# Courage

```{r}
#| label: model
#| cache: true


fit_elec_2 <- brm(formula = death_age ~ treatment + party + win_margin,
                  data = x,
                  family = gaussian(),
                  refresh = 0,
                  silent = 2,
                  seed = 9)

fit_elec_2
```

```{r}
pp_check(fit_elec_2)
tbl_regression(fit_elec_2)
```



# Temprance

## Variables

treatment, win_margin, party

## Values
party: Republican, Democrat, Third Party\n
treatment: win and lose\n
win_margin: 0

## Formula
$$death\_age_i = \beta_{0} + \beta_{1} treatment_i + \beta_{2}party_i + \beta_{3}win\_margin_i + \epsilon_{i}$$

```{r}

treatment <- c("win", "lose")
party <- c("Democrat", "Republican", "Third party")
win_margin <- 0


ndata <- expand_grid(treatment, party, win_margin)
fit_elec_2 |>
  add_epred_draws(newdata = ndata) |>
  ggplot(aes(x = .epred, y = fct_reorder(party, .epred))) +
  stat_slab(aes(fill = treatment), alpha = 0.5) +
  labs(title = "Expected Lifespan Given Election Outcome",
       subtitle = "Winning is associated with about 8 extra years",
       x = "Expected Lifespan",
       y = "",
       fill = "Election Result")
```

